Perceived value of video games, but not hours played, predicts mental well-being in adult Nintendo players

### 2024-07-10 ms.qmd
linting

  - name: Andrew K Przybylski
    corresponding: true
    email: andy.pryzbylski@oii.ox.ac.uk
    orcid: 0000-0003-4126-0696
    affiliations:
      - name: University of Oxford
        department: Oxford Internet Institute
    # roles:
    #   - conceptualization
    #   - funding acquisition
    #   - project administration
    #   - editing
abstract: "Studies on video games and well-being often rely on self-report measures or data from a single game. Here, we study how 704 US adults' time spent playing over 150 1st-party Nintendo Switch games (totaling 140k hours) relates to their life satisfaction, affect, depressive symptoms, and general mental well-being. We replicate previous findings that playtime over the past 2 weeks does not predict well-being, and extend these findings to a wider range of timescales (1 hour to 1 year). Results suggest that relationships, if present, dissipate within 2 hours of gameplay. Our non-causal findings suggest substantial confounding would be needed to shift a meaningful true effect to the observed null. Although playtime was not related to well-being, players’ assessments of the value of game time—so called gaming life fit—was. Results emphasize the importance of defining the gaming population of interest, collecting data from more than one game, and focusing on how players integrate gaming into their lives rather than the amount of time spent."
keywords: [video games, digital trace data, well-being]
author-note:
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    study-registration: This study was registered on the Open Science Framework (https://osf.io/sjqyt).
    # Acknowledge and cite data/materials to be shared.
    data-sharing: Data and materials are available at https://osf.io/6xkdg/. 
    # Example: [Author name] has been a paid consultant for Corporation X, which funded this study.
    conflict-of-interest: Data for this study was provided by Nintendo of America. Nintendo of America had no role in the design of the study, the analysis of the data, or decision to publish any results. The authors declare no other competing interests. 
    # Example: This study was supported by Grant [Grant Number] from [Funding Source].
    financial-support: This research was supported by Huo Family Foundation and the UK Economic and Social Research Council (ES/W012626/1). KM was supported by Forskningsrå det för hälsa, arbetsliv och välfärd (2021-01284).
    # Example: The authors are grateful to [Person] for [Reason].
    gratitude: ~
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: ~
bibliography: references.bib
floatsintext: true # If true, tables and figures are mingled with the text instead of listed at the end of the document.
numbered-lines: false # Numbered lines (.pdf and .docx only)
meta-analysis: false
suppress-credit-statement: true
format:
  apaquarto-docx: default
    # blank-lines-above-author-note: 0
    # blank-lines-above-title: 1
  # apaquarto-html:
  #   documentmode: man
  #   code-fold: true
  #   code-summary: "Show the code"
  #   wordcount-html: default
  # apaquarto-pdf:
  #   documentmode: jou
---

@@

datSurveyRaw <- read_csv("data/survey.csv.gz")
datRaw <- read_csv("data/telemetry.csv.gz") |> 
  left_join(datSurveyRaw, by = "pid")
datDemo <- read_csv("data/demographics.csv.gz")

```

@@

# Function for calculating playtime, relying on vectorized pmin()/pmax()
# When using the durations from Nintendo - we take the duration of sessions that begin or end during the playtime window, and weight them in case of partial sessions 
calculatePlaytime <- function(sessionStart, 
                              sessionEnd, 
                              recordedDate, 
                              duration, # column containing duration as calculated by nintendo
                              timeWindows, 
                              nintendo = FALSE # calculate duration based solely on start/end times (FALSE), or nintendo's heuristics (TRUE)
) {

  overlaps <- map(timeWindows, ~{
    overlapStart <- pmax(sessionStart, as_datetime(recordedDate) %m-% .x)
    overlapEnd <- pmin(sessionEnd, recordedDate)
    overlapTime <- interval(overlapStart, overlapEnd) / dhours(1)

    if (nintendo) {
      weight <- overlapTime/ (interval(sessionStart, sessionEnd) / dhours(1)) 
      overlap <- pmax(0, duration*weight) # if the overlap is negative, return 0
    } else {
      overlap <- pmax(0, overlapTime) # if the overlap is negative, return 0
    }

  })
  # Set the names of the list elements to correspond to the windows
  names(overlaps) <- if (nintendo) { paste0("overlapNintendo", names(timeWindows))} else { paste0("overlap", names(timeWindows)) }
  return(overlaps)
}

mytheme <- list(
  theme_bw() +
    theme(
      strip.background = element_rect(fill = "black"),
      strip.text = element_text(color = "white", size = 10),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      axis.text.y = element_text(size = 10, hjust = .5, vjust = -.015),
      axis.text.x = element_text(size = 10),
      axis.title.x = element_text(hjust = .45, size = 10, face = "bold"),
      axis.title.y = element_text(hjust = .45, size = 10, face = "bold"),
      legend.direction = "vertical",
      legend.text = element_text(size = 10)
    ),
  # coord_cartesian(clip = "off"),
  guides(alpha = "none", fill = guide_legend(reverse = TRUE))
)

@@

nintendoDict <- c(
  "1H" = "playtimeNintendo1hr", "2H" = "playtimeNintendo2hr","6H" = "playtimeNintendo6hr", "12H" = "playtimeNintendo12hr", 
  "1D" = "playtimeNintendo1d", "3D" = "playtimeNintendo3d", "1W" = "playtimeNintendo1wk", "2W" = "playtimeNintendo2wk",
  "1M" = "playtimeNintendo1mo", "3M" = "playtimeNintendo3mo", "6M" = "playtimeNintendo6mo", "1 Year" = "playtimeNintendo1yr"
)

playedDict <- c(
  "1H" = "played1hr", "2H" = "played2hr", "6H" = "played6hr", "12H" = "played12hr",
  "1D" = "played1d", "3D" = "played3d", "1W" = "played1wk", "2W" = "played2wk",
  "1M" = "played1mo", "3M" = "played3mo", "6M" = "played6mo", "1Y" = "played1yr"
)

```

@@

datAll <- datRaw |> 
  lazy_dt() |>
  # Add the time overlaps to the df 
  mutate(across(everything(), ~.x), !!!overlaps,
         across(everything(), ~.x), !!!overlapsNintendo) |> 
  mutate(
    impliedDuration = interval(sessionStart, sessionEnd) / dhours(1)
  ) |> 
  filter(duration < 24*60 & impliedDuration < 24) |> # drop implausible sessions lasting more than 24 hours
  group_by(pid) |>
  summarise(
    across(-starts_with("overlap"), first), # all survey columns are identical within participants, so we just take the first

@@

    # Those longer than one day are divided by the number of days in the interval to indicate
    # mean hours of play per day during that period
    playtime3d = sum(overlap3d)/3,
    playtime1wk = sum(overlap1wk)/7,
    playtime2wk = sum(overlap2wk)/14,
    playtime1mo = sum(overlap1mo)/30,
    playtime3mo = sum(overlap3mo)/90,
    playtime6mo = sum(overlap6mo)/180,
    playtime1yr = sum(overlap1yr)/365,

    # repeat for nintendo duration
    playtimeNintendo1hr = sum(overlapNintendo1hr),
    playtimeNintendo2hr = sum(overlapNintendo2hr),
    playtimeNintendo6hr = sum(overlapNintendo6hr),
    playtimeNintendo12hr = sum(overlapNintendo12hr),
    playtimeNintendo1d = sum(overlapNintendo1d),

    playtimeNintendo3d = sum(overlapNintendo3d)/3,
    playtimeNintendo1wk = sum(overlapNintendo1wk)/7,
    playtimeNintendo2wk = sum(overlapNintendo2wk)/14,
    playtimeNintendo1mo = sum(overlapNintendo1mo)/30,
    playtimeNintendo3mo = sum(overlapNintendo3mo)/90,
    playtimeNintendo6mo = sum(overlapNintendo6mo)/180,
    playtimeNintendo1yr = sum(overlapNintendo1yr)/365,
  ) |>

  # calculate binary versions of each playtime metric (i.e., played at all vs did not play at all)
  mutate(across(starts_with("playtime"),
                ~ . > 0,
                .names = "played{.col}"
         )) |> 
  rename_with(.fn = ~ str_replace(.x, "playtime", ""),
              .cols = starts_with("played")) |>

  select(-(titleID:operationMode), -impliedDuration) |>  # drop telemetry columns; can still be accessed in datRaw
  as_tibble()

@@

datAll <- datAll |>

  mutate(genderRC = ifelse(gender %in% c("Man","Woman",NA), gender, "Non-binary")) |> 
  rowwise() |> 
  mutate(
    meanWemwbs = mean(c_across(starts_with("wemwbs"))),
    meanPromis = mean(c_across(starts_with("promis"))),
    meanLifeFit = mean(c_across(starts_with("displacement"))),
    meanBPS = mean(c_across(starts_with("bps"))),
  ) |> 
  ungroup() |> 
  mutate(
    csas = rescale(csas, from = c(0,10), to = c(1,5)),
    affectiveValence = rescale(affectiveValence_1, from = c(1, 100), to = c(1, 5)),
    .keep = "unused"
    ) %>%

  # Here we run additional checks for potential careless responders using the (1) inter-item correlation and (2) longstring techniques.
  # In the first, we take the within-measure item pairs with the strongest correlation 
  # in the sample, and then identify the participants whose scores substantially differed, suggesting potential carelessness.
  #
  # In the second, we look at the number of consecutive identical answers. 
  # Together, this catches most participants who either gave identical answers to all questions or who chose answers semi-randomly.

  mutate(
    longstring = longstring(select(., bangs_1:bangs_18)),
    intraIndVar = irv(select(., promis_1:promis_8),
                      na.rm = TRUE,
                      num.split = 1)
  ) |> 
  mutate(
    weirdLongstring = longstring > 8,
    weirdPromis = abs(promis_5 - promis_1) >= 4,
    weirdTrojan = abs(trojan_3 - trojan_1) >= 4,
    weirdBangs = abs(bangs_9 - bangs_8) >= 4,
    possibleCareless = ifelse(weirdLongstring | weirdPromis | weirdTrojan | weirdBangs, TRUE, FALSE)
  )
```

@@

# filter out in stages, in case of manual look at change in sample size after each 
dat <- datAll |> 
  filter(!pid %in% nonResponders) |> 
  filter(!pid %in% nonPlayers) |> 
  filter(!pid %in% clockHackers) |> 
  filter(possibleCareless %in% c(FALSE, NA)) |> 
  select(-starts_with(c("wemwbs","bangs","trojan","promis","timeUse","recent","weird")), -c(positives,problematicPlay,feedback))

@@

dat |> 
  left_join(datDemo, by = "pid") |> 
  select(age, gender, employment, eduLevel, ethnicity) |> 
  mutate(
    Age = age,
    Gender = ifelse(!(gender %in% c("Man","Woman")), 
                     "Non-binary or other gender identity", 
                     gender),
    `Employment Status` = ifelse(!(employment %in% c("Working full-time","Working part-time", "Student", "Not currently employed")), 
                     "Other employment status", 
                     employment),
    Ethnicity = ifelse(!(ethnicity %in% c("White","Asian","Mixed","Black")), 
                     "Other ethnicity", 
                     ethnicity),
    `Educational Level` = ifelse(!(eduLevel %in% c("University Bachelors Degree","Asian","Mixed","Black")), 
                     "Other ethnicity", 
                     eduLevel)
  ) |> 
  mutate(
    Gender = fct_relevel(Gender, sort(table(Gender), decreasing = TRUE) %>% names()),
    `Employment Status` = fct_relevel(`Employment Status`, sort(table(`Employment Status`), decreasing = TRUE) %>% names()),
    Ethnicity = fct_relevel(Ethnicity, sort(table(Ethnicity), decreasing = TRUE) %>% names()),
    `Educational Level` = fct_relevel(`Educational Level`, sort(table(`Educational Level`), decreasing = TRUE) %>% names())
  ) |> 
  crosstable(
    cols = c("Age","Gender","Employment Status","Ethnicity"),
    percent_digits = 1,
    funs = c("Mean (SD)" = meansd, "Median (IQR)" = mediqr, "Min / Max" = minmax)
  ) |> 
  crosstable::as_flextable() |> 
  set_header_labels(
    values = c(
      "Descriptor", "Variable", "Value")
  ) |>
  theme_apa() |> 
  fontsize(size = 8, part = "all") |>  # Adjust font size
  padding(padding = 0, part = "all")  # Adjust padding
```

@@

playtime <- median_qi(dat, playtime2wk*14, .width = .7, na.rm = TRUE) |> 
  mutate(across(where(is.numeric), ~round(., 2)))
playtime <- str_glue("{playtime[1]} ({playtime[2]}, {playtime[3]}; 2nd and 8th deciles)")

@@

p1 <- dat |> 
  select(pid, matches("playtime\\d")) |> 
  pivot_longer(-pid) |> 
  mutate(name = fct_rev(fct_recode(name, !!!playtimeDict))) |> 
  mutate(name = fct_inorder(name)) |> 
  ggplot(aes(value, name)) +
  geom_point(alpha = .6) +
  geom_hline(yintercept = 5.5, linetype = "dashed") +
  stat_summary(fun = mean, col = "red") +
  labs(
    x = "Hours of (daily) play",
    y = "Observation Period"
  ) +
  scale_x_continuous(limits = c(0, 6), 
                     breaks = c(0, 1, 2, 3, 4, 5, 6)) +
  mytheme

@@

(p1 | p2) + plot_layout(axes = "collect") + plot_annotation(tag_level = "A")

```

@@

dat |> 
  pivot_longer(cols = c("csas", "affectiveValence", "meanPromis", "meanWemwbs")) |>
  mutate(name = case_when(
    name == "csas" ~ "Life Satisfaction",
    name == "affectiveValence" ~ "Affect",
    name == "meanPromis" ~ "Depressive Symptoms",
    name == "meanWemwbs" ~ "General well-being",
    TRUE ~ NA_character_
    )) |> 
  ggplot(aes(x = playtime2wk*60, y = value)) +
  geom_point(size = .6) +
  geom_smooth(method='lm', formula = y~x, level = .99, color = "darkgreen", fill = "darkgrey") +
  scale_x_sqrt(breaks = c(0, 15, 60, 120, 240, 420)) +
  labs(x = "Minutes of play per day\nduring previous 2 weeks (Square Root Axis)", y = "Outcome (1-5 scale)") +
  mytheme +
  # theme(
  #   axis.line.y = element_line(linewidth = .5)
  # ) + 
  # scale_y_continuous(labels = scales::percent) +
  facet_wrap(~name)

@@

models <- expand.grid(predictor = dat |> select(matches("playtime\\d")) |> names(),
                      outcome = c("csas","affectiveValence","meanWemwbs","meanPromis")) |> 
  mutate(formula = paste0(outcome, " ~ ", predictor, " + age + genderRC + eduLevel + employment"))

fittedModels <- map_df(
  models$formula, ~lm(as.formula(.x), 
                      data = dat) |> 
    broom::tidy(conf.int = TRUE, conf.level = .99) |> 
    filter(grepl("play", term))
) |> 
  mutate(predictor = models$predictor,
         outcome = models$outcome) |> 
  select(predictor, outcome, everything())
```

@@

fittedModels |> 
  mutate(
    model = paste0(outcome, " ~ ", predictor),
    predictor = fct_rev(fct_recode(predictor, !!!playtimeDict)),
    outcome = fct_recode(outcome,
                         "Life Satisfaction" = "csas",
                         "Affect" = "affectiveValence",
                         "Depressive Symptoms" = "meanPromis",
                         "General Mental Well-being" = "meanWemwbs"),

    # some hacky code to match up the time periods with their lubridate::period objects and then convert to numeric
    # so we can plot on the log scale
    window = map(sub("playtime", "", term), ~ timeWindows[.]),
    window = map_dbl(window, ~ as.numeric(.x[[1]], "hours")),
    conf.high.90 = estimate + 1.645*std.error,
    conf.low.90 = estimate - 1.645*std.error
  ) |>
  ggplot(aes(x = window, y = estimate)) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    group = 1,
    alpha = .4,
    fill = "darkgrey"
  ) +
  geom_ribbon(
    aes(ymin = conf.low.90, ymax = conf.high.90),
    group = 1,
    alpha = .8,
    fill = "darkgrey"
  ) +
  geom_line(
    group = 1,
    color = "darkgreen",
    linewidth = 1,
  ) +
  geom_point(color = "darkgreen", size = 2) +
  scale_x_log10(breaks = map_dbl(timeWindows, ~ as.numeric(.x[[1]], "hours")), labels = names(playtimeDict)) +
  labs(
    x = "Observation period for video game play",
    y = "Playtime-outcome association\nin observation period"
  ) +
  theme_linedraw() +
  theme(
    # axis.title.y = element_text(angle = 0, vjust = .5),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(linewidth = .1)
    ) +
  geom_hline(yintercept = .06, linetype = "dashed", linewidth = .3, color = "black") +
  geom_hline(yintercept = -.06, linetype = "dashed", linewidth = .3, color = "black") +
  facet_wrap(~outcome, scales = "free_y")

```

@@

moderationModels <- expand.grid(predictor = dat |> select(matches("playtime\\d")) |> names(),
                      outcome = c("csas","affectiveValence","meanWemwbs","meanPromis")) |> 
  mutate(formula = paste0(outcome, " ~ ", predictor, "*meanLifeFit + ", predictor, "*age + ", predictor, "*genderRC"))

moderationFits <- map_df(
  moderationModels$formula, ~lm(as.formula(.x), 
                                data = dat) |> 
    broom::tidy(conf.int = TRUE, conf.level = .99)
) |> 
  mutate(predictor = rep(moderationModels$predictor, each = 10),
         outcome = rep(moderationModels$outcome, 10)) |> 
  select(predictor, outcome, everything())

directEsts <- moderationFits |> 
  filter(!grepl("playtime", term) & grepl("meanLifeFit", term)) |>
  # filter(!grepl(":", term)) |> 
  mutate(across(estimate, abs),
         across(where(is.numeric), ~round(., 3)))

>>>>

  - name: Andrew K Przybylski
    corresponding: true
    email: andy.pryzbylski@oii.ox.ac.uk
    orcid: 0000-0003-4126-0696
    affiliations:
      - name: University of Oxford
        department: Oxford Internet Institute
    # roles:
    #   - conceptualization
    #   - funding acquisition
    #   - project administration
    #   - editing
abstract: "Studies on video games and well-being often rely on self-report measures or data from a single game. Here, we study how 703 US adults' time spent playing over 150 1st-party Nintendo Switch games (totaling 140k hours) relates to their life satisfaction, affect, depressive symptoms, and general mental well-being. We replicate previous findings that playtime over the past 2 weeks does not predict well-being, and extend these findings to a wider range of timescales (1 hour to 1 year). Results suggest that relationships, if present, dissipate within 2 hours of gameplay. Our non-causal findings suggest substantial confounding would be needed to shift a meaningful true effect to the observed null. Although playtime was not related to well-being, players’ assessments of the value of game time—so called gaming life fit—was. Results emphasize the importance of defining the gaming population of interest, collecting data from more than one game, and focusing on how players integrate gaming into their lives rather than the amount of time spent."
keywords: [video games, digital trace data, well-being]
author-note:
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    study-registration: This study was registered on the Open Science Framework (https://osf.io/sjqyt).
    # Acknowledge and cite data/materials to be shared.
    data-sharing: Data and materials are available at https://osf.io/6xkdg/. 
    # Example: [Author name] has been a paid consultant for Corporation X, which funded this study.
    conflict-of-interest: Data for this study was provided by Nintendo of America. Nintendo of America had no role in the design of the study, the analysis of the data, or decision to publish any results. The authors declare no other competing interests. 
    # Example: This study was supported by Grant [Grant Number] from [Funding Source].
    financial-support: This research was supported by Huo Family Foundation and the UK Economic and Social Research Council (ES/W012626/1). KM was supported by Forskningsrå det för hälsa, arbetsliv och välfärd (2021-01284).
    # Example: The authors are grateful to [Person] for [Reason].
    gratitude: ~
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: ~
bibliography: references.bib
floatsintext: true # If true, tables and figures are mingled with the text instead of listed at the end of the document.
numbered-lines: false # Numbered lines (.pdf and .docx only)
meta-analysis: false
suppress-credit-statement: true
format:
  apaquarto-docx: default
    # blank-lines-above-author-note: 0
    # blank-lines-above-title: 1
  # apaquarto-html:
  #   documentmode: man
  #   code-fold: true
  #   code-summary: "Show the code"
  #   wordcount-html: default
  # apaquarto-pdf:
  #   documentmode: jou
---

@@

datSurveyRaw <- read_csv("data/survey.csv.gz")
datRaw <- read_csv("data/telemetry.csv.gz") |>
  left_join(datSurveyRaw, by = "pid")
datDemo <- read_csv("data/demographics.csv.gz")
```

@@

# Function for calculating playtime, relying on vectorized pmin()/pmax()
# When using the durations from Nintendo - we take the duration of sessions that begin or end during the playtime window, and weight them in case of partial sessions
calculatePlaytime <- function(sessionStart,
                              sessionEnd,
                              recordedDate,
                              duration, # column containing duration as calculated by nintendo
                              timeWindows,
                              nintendo = FALSE # calculate duration based solely on start/end times (FALSE), or nintendo's heuristics (TRUE)
) {
  overlaps <- map(timeWindows, ~ {
    overlapStart <- pmax(sessionStart, as_datetime(recordedDate) %m-% .x)
    overlapEnd <- pmin(sessionEnd, recordedDate)
    overlapTime <- interval(overlapStart, overlapEnd) / dhours(1)

    if (nintendo) {
      weight <- overlapTime / (interval(sessionStart, sessionEnd) / dhours(1))
      overlap <- pmax(0, duration * weight) # if the overlap is negative, return 0
    } else {
      overlap <- pmax(0, overlapTime) # if the overlap is negative, return 0
    }
  })
  # Set the names of the list elements to correspond to the windows
  names(overlaps) <- if (nintendo) {
    paste0("overlapNintendo", names(timeWindows))
  } else {
    paste0("overlap", names(timeWindows))
  }
  return(overlaps)
}

mytheme <- list(
  theme_bw() +
    theme(
      strip.background = element_rect(fill = "black"),
      strip.text = element_text(color = "white", size = 10),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA, size = 1),
      axis.text.y = element_text(size = 10, hjust = .5, vjust = -.015),
      axis.text.x = element_text(size = 10),
      axis.title.x = element_text(hjust = .45, size = 10, face = "bold"),
      axis.title.y = element_text(hjust = .45, size = 10, face = "bold"),
      legend.direction = "vertical",
      legend.text = element_text(size = 10)
    ),
  # coord_cartesian(clip = "off"),
  guides(alpha = "none", fill = guide_legend(reverse = TRUE))
)

@@

nintendoDict <- c(
  "1H" = "playtimeNintendo1hr", "2H" = "playtimeNintendo2hr", "6H" = "playtimeNintendo6hr", "12H" = "playtimeNintendo12hr",
  "1D" = "playtimeNintendo1d", "3D" = "playtimeNintendo3d", "1W" = "playtimeNintendo1wk", "2W" = "playtimeNintendo2wk",
  "1M" = "playtimeNintendo1mo", "3M" = "playtimeNintendo3mo", "6M" = "playtimeNintendo6mo", "1 Year" = "playtimeNintendo1yr"
)

playedDict <- c(
  "1H" = "played1hr", "2H" = "played2hr", "6H" = "played6hr", "12H" = "played12hr",
  "1D" = "played1d", "3D" = "played3d", "1W" = "played1wk", "2W" = "played2wk",
  "1M" = "played1mo", "3M" = "played3mo", "6M" = "played6mo", "1Y" = "played1yr"
)
```

@@

datAll <- datRaw |>
  lazy_dt() |>
  # Add the time overlaps to the df
  mutate(
    across(everything(), ~.x), !!!overlaps,
    across(everything(), ~.x), !!!overlapsNintendo
  ) |>
  mutate(
    impliedDuration = interval(sessionStart, sessionEnd) / dhours(1)
  ) |>
  filter(duration < 24 * 60 & impliedDuration < 24) |> # drop implausible sessions lasting more than 24 hours
  group_by(pid) |>
  summarise(
    across(-starts_with("overlap"), first), # all survey columns are identical within participants, so we just take the first

@@

    # Those longer than one day are divided by the number of days in the interval to indicate
    # mean hours of play per day during that period
    playtime3d = sum(overlap3d) / 3,
    playtime1wk = sum(overlap1wk) / 7,
    playtime2wk = sum(overlap2wk) / 14,
    playtime1mo = sum(overlap1mo) / 30,
    playtime3mo = sum(overlap3mo) / 90,
    playtime6mo = sum(overlap6mo) / 180,
    playtime1yr = sum(overlap1yr) / 365,

    # repeat for nintendo duration
    playtimeNintendo1hr = sum(overlapNintendo1hr),
    playtimeNintendo2hr = sum(overlapNintendo2hr),
    playtimeNintendo6hr = sum(overlapNintendo6hr),
    playtimeNintendo12hr = sum(overlapNintendo12hr),
    playtimeNintendo1d = sum(overlapNintendo1d),
    playtimeNintendo3d = sum(overlapNintendo3d) / 3,
    playtimeNintendo1wk = sum(overlapNintendo1wk) / 7,
    playtimeNintendo2wk = sum(overlapNintendo2wk) / 14,
    playtimeNintendo1mo = sum(overlapNintendo1mo) / 30,
    playtimeNintendo3mo = sum(overlapNintendo3mo) / 90,
    playtimeNintendo6mo = sum(overlapNintendo6mo) / 180,
    playtimeNintendo1yr = sum(overlapNintendo1yr) / 365,
  ) |>
  # calculate binary versions of each playtime metric (i.e., played at all vs did not play at all)
  mutate(across(starts_with("playtime"),
    ~ . > 0,
    .names = "played{.col}"
  )) |>
  rename_with(
    .fn = ~ str_replace(.x, "playtime", ""),
    .cols = starts_with("played")
  ) |>
  select(-(titleID:operationMode), -impliedDuration) |> # drop telemetry columns; can still be accessed in datRaw
  as_tibble()

@@

datAll <- datAll |>
  mutate(genderRC = ifelse(gender %in% c("Man", "Woman", NA), gender, "Non-binary")) |>
  rowwise() |>
  mutate(
    meanWemwbs = mean(c_across(starts_with("wemwbs"))),
    meanPromis = mean(c_across(starts_with("promis"))),
    meanLifeFit = mean(c_across(starts_with("displacement"))),
    meanBPS = mean(c_across(starts_with("bps"))),
  ) |>
  ungroup() |>
  mutate(
    csas = rescale(csas, from = c(0, 10), to = c(1, 5)),
    affectiveValence = rescale(affectiveValence_1, from = c(1, 100), to = c(1, 5)),
    .keep = "unused"
  ) %>%
  # Here we run additional checks for potential careless responders using the (1) inter-item correlation and (2) longstring techniques.
  # In the first, we take the within-measure item pairs with the strongest correlation
  # in the sample, and then identify the participants whose scores substantially differed, suggesting potential carelessness.
  #
  # In the second, we look at the number of consecutive identical answers.
  # Together, this catches most participants who either gave identical answers to all questions or who chose answers semi-randomly.

  mutate(
    longstring = longstring(select(., bangs_1:bangs_18)),
    intraIndVar = irv(select(., promis_1:promis_8),
      na.rm = TRUE,
      num.split = 1
    )
  ) |>
  mutate(
    weirdLongstring = longstring > 8,
    weirdPromis = abs(promis_5 - promis_1) >= 4,
    weirdTrojan = abs(trojan_3 - trojan_1) >= 4,
    weirdBangs = abs(bangs_9 - bangs_8) >= 4,
    possibleCareless = ifelse(weirdLongstring | weirdPromis | weirdTrojan | weirdBangs, TRUE, FALSE)
  )
```

@@

# filter out in stages, in case of manual look at change in sample size after each
dat <- datAll |>
  filter(!pid %in% nonResponders) |>
  filter(!pid %in% nonPlayers) |>
  filter(!pid %in% clockHackers) |>
  filter(possibleCareless %in% c(FALSE, NA)) |>
  select(-starts_with(c("wemwbs", "bangs", "trojan", "promis", "timeUse", "recent", "weird")), -c(positives, problematicPlay, feedback))

@@

dat |>
  left_join(datDemo, by = "pid") |>
  select(age, gender, employment, eduLevel, ethnicity) |>
  mutate(
    Age = age,
    Gender = ifelse(!(gender %in% c("Man", "Woman")),
      "Non-binary or other gender identity",
      gender
    ),
    `Employment Status` = ifelse(!(employment %in% c("Working full-time", "Working part-time", "Student", "Not currently employed")),
      "Other employment status",
      employment
    ),
    Ethnicity = ifelse(!(ethnicity %in% c("White", "Asian", "Mixed", "Black")),
      "Other ethnicity",
      ethnicity
    ),
    `Educational Level` = ifelse(!(eduLevel %in% c("University Bachelors Degree", "Asian", "Mixed", "Black")),
      "Other ethnicity",
      eduLevel
    )
  ) |>
  mutate(
    Gender = fct_relevel(Gender, sort(table(Gender), decreasing = TRUE) %>% names()),
    `Employment Status` = fct_relevel(`Employment Status`, sort(table(`Employment Status`), decreasing = TRUE) %>% names()),
    Ethnicity = fct_relevel(Ethnicity, sort(table(Ethnicity), decreasing = TRUE) %>% names()),
    `Educational Level` = fct_relevel(`Educational Level`, sort(table(`Educational Level`), decreasing = TRUE) %>% names())
  ) |>
  crosstable(
    cols = c("Age", "Gender", "Employment Status", "Ethnicity"),
    percent_digits = 1,
    funs = c("Mean (SD)" = meansd, "Median (IQR)" = mediqr, "Min / Max" = minmax)
  ) |>
  crosstable::as_flextable() |>
  set_header_labels(
    values = c(
      "Descriptor", "Variable", "Value"
    )
  ) |>
  theme_apa() |>
  fontsize(size = 8, part = "all") |> # Adjust font size
  padding(padding = 0, part = "all") # Adjust padding
```

@@

playtime <- median_qi(dat, playtime2wk * 14, .width = .7, na.rm = TRUE) |>
  mutate(across(where(is.numeric), ~ round(., 2)))
playtime <- str_glue("{playtime[1]} ({playtime[2]}, {playtime[3]}; 2nd and 8th deciles)")

@@

p1 <- dat |>
  select(pid, matches("playtime\\d")) |>
  pivot_longer(-pid) |>
  mutate(name = fct_rev(fct_recode(name, !!!playtimeDict))) |>
  mutate(name = fct_inorder(name)) |>
  ggplot(aes(value, name)) +
  geom_point(alpha = .6) +
  geom_hline(yintercept = 5.5, linetype = "dashed") +
  stat_summary(fun = mean, col = "red") +
  labs(
    x = "Hours of (daily) play",
    y = "Observation Period"
  ) +
  scale_x_continuous(
    limits = c(0, 6),
    breaks = c(0, 1, 2, 3, 4, 5, 6)
  ) +
  mytheme

@@

(p1 | p2) + plot_layout(axes = "collect") + plot_annotation(tag_level = "A")
```

@@

dat |>
  pivot_longer(cols = c("csas", "affectiveValence", "meanPromis", "meanWemwbs")) |>
  mutate(name = case_when(
    name == "csas" ~ "Life Satisfaction",
    name == "affectiveValence" ~ "Affect",
    name == "meanPromis" ~ "Depressive Symptoms",
    name == "meanWemwbs" ~ "General well-being",
    TRUE ~ NA_character_
  )) |>
  ggplot(aes(x = playtime2wk * 60, y = value)) +
  geom_point(size = .6) +
  geom_smooth(method = "lm", formula = y ~ x, level = .99, color = "darkgreen", fill = "darkgrey") +
  scale_x_sqrt(breaks = c(0, 15, 60, 120, 240, 420)) +
  labs(x = "Minutes of play per day\nduring previous 2 weeks (Square Root Axis)", y = "Outcome (1-5 scale)") +
  mytheme +
  # theme(
  #   axis.line.y = element_line(linewidth = .5)
  # ) +
  # scale_y_continuous(labels = scales::percent) +
  facet_wrap(~name)

@@

models <- expand.grid(
  predictor = dat |> select(matches("playtime\\d")) |> names(),
  outcome = c("csas", "affectiveValence", "meanWemwbs", "meanPromis")
) |>
  mutate(formula = paste0(outcome, " ~ ", predictor, " + age + genderRC + eduLevel + employment"))

fittedModels <- map_df(
  models$formula, ~ lm(as.formula(.x),
    data = dat
  ) |>
    broom::tidy(conf.int = TRUE, conf.level = .99) |>
    filter(grepl("play", term))
) |>
  mutate(
    predictor = models$predictor,
    outcome = models$outcome
  ) |>
  select(predictor, outcome, everything())
```

@@

fittedModels |>
  mutate(
    model = paste0(outcome, " ~ ", predictor),
    predictor = fct_rev(fct_recode(predictor, !!!playtimeDict)),
    outcome = fct_recode(outcome,
      "Life Satisfaction" = "csas",
      "Affect" = "affectiveValence",
      "Depressive Symptoms" = "meanPromis",
      "General Mental Well-being" = "meanWemwbs"
    ),

    # some hacky code to match up the time periods with their lubridate::period objects and then convert to numeric
    # so we can plot on the log scale
    window = map(sub("playtime", "", term), ~ timeWindows[.]),
    window = map_dbl(window, ~ as.numeric(.x[[1]], "hours")),
    conf.high.90 = estimate + 1.645 * std.error,
    conf.low.90 = estimate - 1.645 * std.error
  ) |>
  ggplot(aes(x = window, y = estimate)) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    group = 1,
    alpha = .4,
    fill = "darkgrey"
  ) +
  geom_ribbon(
    aes(ymin = conf.low.90, ymax = conf.high.90),
    group = 1,
    alpha = .8,
    fill = "darkgrey"
  ) +
  geom_line(
    group = 1,
    color = "darkgreen",
    linewidth = 1,
  ) +
  geom_point(color = "darkgreen", size = 2) +
  scale_x_log10(breaks = map_dbl(timeWindows, ~ as.numeric(.x[[1]], "hours")), labels = names(playtimeDict)) +
  labs(
    x = "Observation period for video game play",
    y = "Playtime-outcome association\nin observation period"
  ) +
  theme_linedraw() +
  theme(
    # axis.title.y = element_text(angle = 0, vjust = .5),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(linewidth = .1)
  ) +
  geom_hline(yintercept = .06, linetype = "dashed", linewidth = .3, color = "black") +
  geom_hline(yintercept = -.06, linetype = "dashed", linewidth = .3, color = "black") +
  facet_wrap(~outcome, scales = "free_y")
```

@@

moderationModels <- expand.grid(
  predictor = dat |> select(matches("playtime\\d")) |> names(),
  outcome = c("csas", "affectiveValence", "meanWemwbs", "meanPromis")
) |>
  mutate(formula = paste0(outcome, " ~ ", predictor, "*meanLifeFit + ", predictor, "*age + ", predictor, "*genderRC"))

moderationFits <- map_df(
  moderationModels$formula, ~ lm(as.formula(.x),
    data = dat
  ) |>
    broom::tidy(conf.int = TRUE, conf.level = .99)
) |>
  mutate(
    predictor = rep(moderationModels$predictor, each = 10),
    outcome = rep(moderationModels$outcome, 10)
  ) |>
  select(predictor, outcome, everything())

directEsts <- moderationFits |>
  filter(!grepl("playtime", term) & grepl("meanLifeFit", term)) |>
  # filter(!grepl(":", term)) |>
  mutate(
    across(estimate, abs),
    across(where(is.numeric), ~ round(., 3))
  )
